Trait plasticity and covariance along a continuous soil moisture gradient

### 2021-01-17 FVT Manuscript/manuscript.Rmd
V5

abstract: Water availality is perhaps the greatest environmental determinant of plant yield and fitness. Yet our understanding of plant-water relations is limited because it is primarily informed by experiments considering soil moisture variability at two discrete levels - wet and dry - rather than as a continually varying environmental gradient. Here we used experimental and statistical methods based on function-valued traits to explore responses to continuously varying soil moisture gradient in physiological and morphological traits in two species and five genotypes each of the model grass \textit{Brachypodium}. We find that most traits exhibit non-linear responses to soil moisture variability. We also observe differences in the shape of these non-linear responses between traits, species, and genotypes. Emergent phenomena arise from this variation including changes in trait correlations and evolutionary constraints as a function of soil moisture. These results point to the importance of considering non-linearity in plant-water relations to understand plastic and evolutionary responses to changing climates.

@@

## Analyses
We used R for all statistical analyses. Code and data to generate this manuscript can be found at https://github.com/greymonroe/brachypodium_fvt.

>>>>

abstract: Water availality is perhaps the greatest environmental determinant of plant yield and fitness. However, our understanding of plant-water relations is limited because it is primarily informed by experiments considering soil moisture variability at two discrete levels - wet and dry - rather than as a continuously varying environmental gradient. Here we used experimental and statistical methods based on function-valued traits to explore responses to a continuous soil moisture gradient in physiological and morphological traits in two species and five genotypes each of the model grass \textit{Brachypodium}. We find that most traits exhibit non-linear responses to soil moisture variability. We also observe differences in the shape of these non-linear responses between traits, species, and genotypes. Emergent phenomena arise from this variation including changes in trait correlations and evolutionary constraints as a function of soil moisture. These results point to the importance of considering non-linearity in plant-water relations to understand plastic and evolutionary responses to changing climates.

@@

## Analyses
We used R for all statistical analyses. Code and data to generate this manuscript can be found at https://github.com/greymonroe/brachypodium_fvt and Zenodo (10.5281/zenodo.4446263).

### 2021-01-17 FVT Manuscript/manuscript.tex
V5

\abstract{
Water availality is perhaps the greatest environmental determinant of plant yield and fitness. Yet our understanding of plant-water relations is limited because it is primarily informed by experiments considering soil moisture variability at two discrete levels - wet and dry - rather than as a continually varying environmental gradient. Here we used experimental and statistical methods based on function-valued traits to explore responses to continuously varying soil moisture gradient in physiological and morphological traits in two species and five genotypes each of the model grass \textit{Brachypodium}. We find that most traits exhibit non-linear responses to soil moisture variability. We also observe differences in the shape of these non-linear responses between traits, species, and genotypes. Emergent phenomena arise from this variation including changes in trait correlations and evolutionary constraints as a function of soil moisture. These results point to the importance of considering non-linearity in plant-water relations to understand plastic and evolutionary responses to changing climates.

@@

We used R for all statistical analyses. Code and data to generate this manuscript can be found at \url{https://github.com/greymonroe/brachypodium_fvt}.

>>>>

\abstract{
Water availality is perhaps the greatest environmental determinant of plant yield and fitness. However, our understanding of plant-water relations is limited because it is primarily informed by experiments considering soil moisture variability at two discrete levels - wet and dry - rather than as a continuously varying environmental gradient. Here we used experimental and statistical methods based on function-valued traits to explore responses to a continuous soil moisture gradient in physiological and morphological traits in two species and five genotypes each of the model grass \textit{Brachypodium}. We find that most traits exhibit non-linear responses to soil moisture variability. We also observe differences in the shape of these non-linear responses between traits, species, and genotypes. Emergent phenomena arise from this variation including changes in trait correlations and evolutionary constraints as a function of soil moisture. These results point to the importance of considering non-linearity in plant-water relations to understand plastic and evolutionary responses to changing climates.

@@

We used R for all statistical analyses. Code and data to generate this manuscript can be found at \url{https://github.com/greymonroe/brachypodium_fvt} and Zenodo (10.5281/zenodo.4446263).
